Synthesizing Symmetric Lenses

### 2018-07-08 papers/SymmetricLenses/main.tex
stuff

% begin synthesis
\section{Synthesis}
\label{sec:synthesis}
Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm. The
input regular expressions are first converted into stochastic regular
expressions with \ToStochastic. Next, $\RXSearch$ is initialized with that pair
of stochastic regular expressions. $\RXSearch$ then proposes candidate classes
of lenses by proposing regular expression pairs that characterize those classes.
These lens classes are proposed based on the distance from the original
stochastic regular expressions: classes characterized by stochastic regular
expressions closer to the originals are proposed earlier. $\GreedySynth$
greedily finds a low cost lens within each proposed lens class, if there is one
that satisfies the specification. These lenses are then provided to $\RXSearch$.
Eventually, $\RXSearch$ terminates and returns the best lens found thus far.

\begin{algorithm}
  \caption{\SynthSymLens}
  \label{alg:synth-sym-lens}
  \begin{algorithmic}[1]
    \Function{SynthSymLens}{$\Regex,\RegexAlt,\Examples$}
    \State $\Regex \gets \Call{\ToStochastic}{\Regex}$
    \State $\RegexAlt \gets \Call{\ToStochastic}{\RegexAlt}$
    \State $\RXSearchState \gets \Call{\RXSearch.Init}{\Regex,\RegexAlt}$
    \While{$\Call{\RXSearch.Continue}{\RXSearchState}$}
    \State $(\Regex,\RegexAlt,\RXSearchState) \gets \Call{\RXSearch.Propose}{\RXSearchState}$
    \State $\LCO \gets \Call{\GreedySynth}{\Examples,\Regex,\RegexAlt}$
    \State $\RXSearchState \gets \Call{\RXSearch.Consider}{\RXSearchState,\LCO}$
    \EndWhile
    \State \ReturnVal{\Call{RXSearch.Best}{\RXSearchState}}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\subsection{\RXSearch}
The first synthesizer is one that searches through regular expression pairs.
This synthesizer, \RXSearch, proposes regular expressions, semantically but not
syntactically equivalent to the originals.  In particular, this synthesizer
performs the following rewrites:
\begin{center}
  \begin{tabular}{rcl}
    $\PRegexStar{\Regex}{\Probability}$
    & \Rewrite
    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability}\\

@@

\RXSearch terminates when the sum of the number of regular expressions of the
smallest unproposed size with the distance of those regular expressions from the
originals exceeds the cost of the current best lens. At this point, \RXSearch
stops proposing regular expression pairs, and instead returns to the user the
best lens found thus far. Note this means that if \RXSearch finds a bijective
lens, which has zero cost it will immediately return.
%
\subsection{Stochastic DNF Regular Expressions}
Stochastic DNF regular expressions are a highly normalized form of stochastic
regular expressions. Due to this highly normalized nature, where type
equivalence is often needed for simple symmetric lenses, such rules aren't
needed when typed between DNF regular expressions, as many semantically equal
but syntactically distinct regular expressions have syntactically equal DNF
forms.  Like how stochastic regular expressions are regular expressions with
probability annotations, stochastic DNF regular expressions are DNF regular
expressions with probability annotations.  Like with SREs and regular
expressions, if $\DNFRegex$ is a stochastic DNF regular expression,
$\overline{\DNFRegex}$ is the DNF regular expression generated by erasing
probability annotations.

>>>>

% begin synthesis
\section{Synthesis}
\label{sec:synthesis}
Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm at a
high level of abstraction. The
input regular expressions are first converted into stochastic regular
expressions with \ToStochastic. Once that conversion has taken place,
the $\RXSearch$ algorithm executes.
This algorithm itself is defined via
five functions:
(1) $\PCF{\RXSearch.Create}$,
(2) $\PCF{\RXSearch.Continue}$,
(3) $\PCF{\RXSearch.Propose}$,
(4) $\PCF{\RXSearch.Enqueue}$, and
(5) $\PCF{\RXSearch.Best}$.
First, $\PCF{\RXSearch.Create}$ creates a priority queue ($rxs$)
containing a single element:  the initial pair of stochastic regular
expressions.  Next, 
$\RXSearch$ enters a loop that searches for likely lenses.  The loop
terminates when the algorithm believes it is unlikely
to find a better lens than the best one it has found so far (a
termination condition defined by $\PCF{\RXSearch.Continue}$).
Within each iteration of the loop, it:
\begin{itemize}
\item generates the next class ($S$, $T$) of lenses to
search ($\PCF{\RXSearch.Propose}$),
\item executes $\GreedySynth$ to find a best lens in that class if
one exists ($lco$),
using the examples $exs$ to filter out potential lenses that do not satisfy
the specification, and
\item adds this best lens to the set of possibilities
under consideration $\PCF{\RXSearch.Add}$
\end{itemize}
When the loop terminates, the search returns the globally best lens found
($\PCF{\RXSearch.Best}$).  Each element of this algorithm will be explained in
further depth in the following subsections.

\dpw{Note: I added overbars to regexps in Algorithm 1 below ... double check me
for correctness.}
\begin{algorithm}
  \caption{\SynthSymLens}
  \label{alg:synth-sym-lens}
  \begin{algorithmic}[1]
    \Function{SynthSymLens}{$\BRegex,\BRegexAlt,\Examples$}
    \State $\Regex \gets \Call{\ToStochastic}{\BRegex}$
    \State $\RegexAlt \gets \Call{\ToStochastic}{\BRegexAlt}$
    \State $\RXSearchState \gets \Call{\RXSearch.Create}{\Regex,\RegexAlt}$
    \While{$\Call{\RXSearch.Continue}{\RXSearchState}$}
    \State $(\Regex,\RegexAlt,\RXSearchState) \gets \Call{\RXSearch.Propose}{\RXSearchState}$
    \State $\LCO \gets \Call{\GreedySynth}{\Examples,\Regex,\RegexAlt}$
    \State $\RXSearchState \gets \Call{\RXSearch.Add}{\RXSearchState,\LCO}$
    \EndWhile
    \State \ReturnVal{\Call{RXSearch.Best}{\RXSearchState}}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\subsection{\PCF{\RXSearch.Propose}}
The first synthesizer is one that searches through regular expression pairs.
This synthesizer, \RXSearch, proposes regular expressions, semantically but not
syntactically equivalent to the originals.  In particular, this synthesizer
performs the following rewrites:
\begin{center}
  \begin{tabular}{rcl}
    $\PRegexStar{\Regex}{\Probability}$
    & \Rewrite
    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability}\\

@@

\RXSearch terminates when the sum of the number of regular expressions of the
smallest unproposed size with the distance of those regular expressions from the
originals exceeds the cost of the current best lens. At this point, \RXSearch
stops proposing regular expression pairs, and instead returns to the user the
best lens found thus far. Note this means that if \RXSearch finds a bijective
lens, which has zero cost it will immediately return.
%
\subsection{Stochastic DNF Regular Expressions}
Stochastic DNF regular expressions (\SDNFREabbrev) are a pseudo-normal form
for stochastic regular expressions.  Of all the star-semiring equations
required to relate two \SDNFREabbrevs, only those relating to Kleene star
(\EG, \emph{Unrollstar$_L$} and \emph{Unrollstar$_R$}) are needed to relate
two \SDNFREabbrevs, other axioms, like associativity, commutativity
or distributivity, are unneeded.  This important property makes it possible to
construct an efficient algorithm ($\GreedySynth$), directed
by the syntax of two \SDNFREabbrevs, to synthesize any lenses in the
same ``class'' (\IE, any lenses that have the same type when the
Kleene star axioms are unused).  This property, and the resulting
efficiency of $\GreedySynth$ is key to the effectiveness of our
core algorithms.

Just as SREs are REs with probability annotations,
\SDNFREabbrevs are \DNFREabbrevs with probability annotations.
As with SREs and regular
expressions, if $\DNFRegex$ is a stochastic DNF regular expression,
$\overline{\DNFRegex}$ is the DNF regular expression generated by erasing
probability annotations.
